---
title: "Example Injury Surveillance Report"
author: "Dr Sean Williams"
params:
  season: 2023-24
output:
  rmarkdown::html_document:
    theme: paper
    highlight: monochrome
    toc: yes
    toc_depth: 3
    number_sections: yes
    fig_caption: yes
    toc_float:
      collapsed: no
      smooth_scroll: yes
    always_allow_html: yes
---

```{r results=FALSE, message=FALSE, warning=FALSE, include=FALSE}

# load libraries - these need to be installed first

library(readxl)
library(tidyverse)
library(janitor)
library(stringr)
library(ggrepel)
library(DT)
library(here)

# set theme for plots
theme_set(theme_bw())

# Function to convert all text to sentence case
to_sentence_case <- function(x) {
  str_to_sentence()
}

# Set the working directory using here()
here::here()

# Load and clean injury data for current season 

injuries <- read_excel(here("data", "example_match_injury_data.xlsx")) %>% 
  clean_names() %>% 
  clean_names() %>% 
  mutate(days_absent = as.numeric(days_absent)) %>% # days absent to numeric
  mutate(across(where(is.character), str_to_sentence)) %>% # make all text sentence case
  mutate_if(is.character, as.factor) 

# set overall match exposure time

match_exposure <- read_excel(here("data", "example_match_injury_data.xlsx"), sheet = "Exposure data") %>%
  clean_names() %>% 
  summarise(exposure_time = sum(match_exposure))

match_exposure <-  match_exposure$exposure_time

# overall injury descriptives

match_summary  <- injuries %>%
  summarise(
    n = n(),
    rate = round((n / match_exposure * 1000), 0),
    severity_mean = round(mean(days_absent, na.rm = T), 0),
    severity_median = round(median(days_absent, na.rm = T), 0),
    burden = round((rate * severity_mean), 0))

# injuries as a function of body region

body_region_summary  <- injuries %>%
  group_by(location_main) %>% 
  summarise(
    n = n(),
    rate = round((n / match_exposure * 1000), 0),
    severity_mean = round(mean(days_absent, na.rm = T), 0),
    severity_median = round(median(days_absent, na.rm = T), 0),
    burden = round((rate * severity_mean), 0)) %>% 
  arrange(desc(burden))
  
# injuries as a function of body location

location_summary  <- injuries %>%
  group_by(location) %>% 
  summarise(
    n = n(),
    rate = round((n / match_exposure * 1000), 0),
    severity_mean = round(mean(days_absent, na.rm = T), 0),
    severity_median = round(median(days_absent, na.rm = T), 0),
    burden = round((rate * severity_mean), 0)) %>% 
  arrange(desc(burden))

```

![Photo by [Olga Guryanova](https://unsplash.com/@designer4u) on [Unsplash](https://unsplash.com/photos/men-playing-football-ft7vJxwl2RY?utm_content=creditCopyText&utm_medium=referral&utm_source=unsplash).](images/report_cover_page.jpg)

# Introduction

Welcome to the `r params$season` season report. This injury surveillance project plays a pivotal role in both providing the baseline data needed to assess trends in injury risk, and in guiding further investigation into injuries that are common, severe, or increasing in incidence.

## Overall injury risk

In the `r params$season` season there were **`r match_summary[1,1]` auditable match injuries** that led to time-loss from training and/or match play. The analysis included `r match_exposure` total hours of match exposure, resulting in a match injury incidence rate of **`r match_summary[1,2]` injuries per 1000 match‐hours**. The mean number of days lost per injury during the `r params$season` season was **`r match_summary[1,3]` days**, and the median severity was **`r match_summary[1,4]` days**. The overall injury burden was **`r match_summary[1,5]` days per 1000 match‐hours**.

### Body region

A risk matrix for injuries as a function of body region is presented below. `r body_region_summary[1,1]` injuries had the highest overall injury burden (`r body_region_summary[1,5]` days per 1000 match-hours).

```{r message = FALSE, echo=FALSE, warning=FALSE}

ggplot(body_region_summary, aes(x = rate, y = severity_mean, size = burden, label = location_main)) +
  geom_point(alpha = 0.7) +
  geom_label_repel(
    size = 3, 
    box.padding = 0.5, 
    point.padding = 0.5, 
    segment.color = "grey"
  ) +
  labs(title = "",
       y = "Mean days missed per injury",
       x = "Injury incidence rate (per 1000 match-hours)",
       size = "Injury burden")
```

### Injury location

A breakdown of injuries as a function of body location is presented in the table below. The injury locations with the highest injury burden were 1) `r location_summary[1,1]`, 2) `r location_summary[2,1]`, and 3) `r location_summary[3,1]`.

```{r message = FALSE, echo=FALSE, warning=FALSE}
datatable(
  location_summary,
  colnames = c(
    "Location",
    "Injury count",
    "Incidence rate (injuries / 1000 h)",
    "Mean severity",
    "Median severity",
    "Burden (days missed / 1000 h) "
  )
) 
```
